Modern Graph Generators: A Survey

### 2018-09-24 challenges.tex
Sherif Changes

\subsection{Simple Usage, Simple Parameters}
The proposal of a data generator (not necessarily for graph data) has to face an
important schism. On one hand, it must provide the user with as
many parameters as possible in order to enable him/her to generate
arbitrary data.
This approach seems to be reasonable, but it entails a shortcoming due to
the fact that ordinary users are
unwilling to use complex benchmarking tools. This observation can be
seen, for example, in the case of XML benchmarks -- even though there exist
robust and complex data generators (such as
ToXGene~\cite{conf/webdb/BarbosaMKL02}, which supports specification of
structural aspects, value distributions, references etc.), the most popular
benchmarking tool is XMark~\cite{Schmidt:2002:XBX:1287369.1287455}, which models
a single use case and enables to specify just the size of the data. Hence, the
other extreme is to provide a simple data generator which does not require any
complex settings and thus guarantees a simple and fast benchmarking process.

@@

Most of existing graph generators aiming at generating large graphs with realistic
structural characteristics focus principally on reproducing the degree
distribution and the clustering
coefficient~\cite{kolda2014scalable,edunov2016darwini}. However, there are other
structural characteristics one might be interested in reproducing for a large
graph, such as the diameter, the size of the largest connected component, or the
hierarchical community structure. Graph practitioners are highly interested in knowing
how other high-level structural characteristics affect the performance of graph
queries and graph algorithms. Hence, a compelling open challenge consists
of creating
graph generators that allow to reproduce diverse structural characteristics
of the graphs along with large scale sizes.

@@

Most of existing graph generators generate graphs that are either not labeled or
are specific to a given domain (e.g.  Social Networks). Graphs from different
domains have different schemas, structural characteristics, property
distributions, etc. which might have an impact on the performance of the
application under test. Thus, graph processing engine developers are asking for
generators or tools to flexibly and holistically generate multi-domain graphs, allowing to configure aspects
such as size, schemata, data distributions and other structural
characteristics such as degree distributions, clustering coefficients, and
so on.

@@

The evolution of graphs can be considered variously. We can assume a static set of nodes and a varying set of edges.  Or, there are applications where the graphs only ``grow'', i.e., the set of nodes and/or edges is only extended with new items. In the most general case we can assume any changes in both set of nodes and set of edges. Anyway with the evolution aspect the complexity of classical graph problems increases significantly~\cite{Michail2015,Wu:2014:PPT:2732939.2732945}. In some graph applications, such as, e.g., social networks, the evolution of the
data is a significant aspect, especially in the activity graphs~\cite{doreian1997evolution,Kumar:2006:SEO:1150402.1150476,Hellmann2014583,wang2013,Kossinets88,Viswanath:2009:EUI:1592665.1592675}.
However, as shown in~\cite{Leskovec:2005:RMT:2101235.2101254,Leskovec:2005:GOT:1081870.1081893},
evolving graphs have further specific features. For example, some graphs grow over time according to a \emph{densification power law} which means that real graphs tend to sprout many more edges than nodes, and thus are densifying as they grow. Also the effective diameter of graphs tends to shrink or stabilize as the graph grows with time.

A related problem is \emph{data versioning} and respective ability to query across multiple versions of data or to carry out general analysis.
%Considering graph
%data this problem is also common, for example  in the area of Linked
This problem has been investigated for instance within the domain of Linked
Open Data~\cite{DBLP:conf/semweb/Papakonstantinou16,DBLP:conf/esws/MeimarisP16,fernandez2015towards,fernandez2015bear}.

The respective data generator should hence be able to simulate a natural growth and/or changes in the structure of the graph with regards to the various features of distinct use cases. However,  even though the area of dynamic graphs is intensively studied, surprisingly there seem to exist only very few proposals of a generator dealing with this area. In~\cite{GoerkeKlugeSchumm2012_1000029825} the authors focus on \emph{clustering dynamic graphs}, i.e. graphs where the clustering corresponds to the partition of nodes into natural groups based on the paradigm of intra-cluster density versus inter-cluster sparsity of edges. The generator generates a time series of random graphs $G_0, G_1, ..., G_n$, where $G_t$ emerges from $G_{t-1}$ via atomic updates, i.e., insertion or deletion of an edge or a vertex. The generator keeps track of a (dynamic) ground truth clustering. The probability of atomic events is chosen in a way that adheres to this clustering, without losing randomness.

@@

\subsection{Multi-model Data}
With the dawn of Big Data and especially its variety aspect, new types of
database management systems have emerged. One of the most interesting ones
are the so-called \emph{multi-model databases} that enable to store and
thus query across structurally different data. There exist various types of
multi-model systems combining  distinct subsets of Big Data structures including graph data.
For example, OrientDB\footnote{\url{http://orientdb.com/orientdb/}} which was
implemented on the basis of an object DBMS currently supports graph, document,
key/value, and object models. Such type of DBMSs also needs a specific
data generator that would enable
to test new features and analyze efficiency of operations. However, since the
multi-model systems are in the context of Big Data rather new, there exist only
a few benchmarks targeting multi-model DBMSs (such as
Bigframe~\cite{journals/pvldb/KunjirKB14} or UniBench~\cite{conf/cidr/lu17})
with limited capabilities.

>>>>

\subsection{Simple Usage, Simple Parameters}
The proposal of a data generator (not necessarily for graph data) has to face an
important schism. On one hand, it must provide the user with as
many parameters as possible in order to enable him/her to generate
arbitrary data.
This approach seems to be reasonable, but it entails a shortcoming due to
the fact that ordinary users are
unwilling to use complex benchmarking tools. This observation can be
seen, for example, in the case of XML benchmarks -- even though there exist
robust and complex data generators (such as
ToXGene~\cite{conf/webdb/BarbosaMKL02}, which supports the specification of
structural aspects, value distributions, references etc.), the most popular
benchmarking tool is XMark~\cite{Schmidt:2002:XBX:1287369.1287455}, which models
a single use case and enables its users to specify just the size of the data. Hence, the
other extreme is to provide a simple data generator which does not require any
complex settings and thus guarantees a simple and fast benchmarking process.

@@

Most of existing graph generators are focused on generating large graphs with realistic
structural characteristics focus principally on reproducing the degree
distribution and the clustering
coefficient~\cite{kolda2014scalable,edunov2016darwini}. However, there are other
structural characteristics that one might be interested in reproducing for a large
graph, such as the diameter, the size of the largest connected component, or the
hierarchical community structure. Graph practitioners are highly interested in knowing
how other high-level structural characteristics affect the performance of graph
queries and graph algorithms. Hence, a compelling open challenge consists
of creating
graph generators that allow one to reproduce diverse structural characteristics
of the graphs along with large scale sizes.

@@

Most of existing graph generators also generate graphs that are either not labeled or
are specific to a given domain (e.g.  Social Networks). Graphs from different
domains have different schemas, structural characteristics, property
distributions, etc. which might have an impact on the performance of the
application under test. Thus, graph processing engine developers are asking for
generators or tools to flexibly and holistically generate multi-domain graphs  in a flexible and holistic manner, allowing to configure aspects
such as size, schemata, data distributions and other structural
characteristics such as degree distributions, clustering coefficients, and
so on.

@@

The evolution of graphs can be considered from multiple perspectives. We can assume a static set of nodes and a varying set of edges.  Or, there are applications where the graphs only ``grow'', i.e., the set of nodes and/or edges is only extended with new items. In the most general case we can assume any changes in both set of nodes and set of edges. Anyway with the evolution aspect the complexity of classical graph problems increases significantly~\cite{Michail2015,Wu:2014:PPT:2732939.2732945}. In some graph applications, such as, e.g., social networks, the evolution of the
data is a significant aspect, especially in the activity graphs~\cite{doreian1997evolution,Kumar:2006:SEO:1150402.1150476,Hellmann2014583,wang2013,Kossinets88,Viswanath:2009:EUI:1592665.1592675}.
However, as shown in~\cite{Leskovec:2005:RMT:2101235.2101254,Leskovec:2005:GOT:1081870.1081893},
evolving graphs have further specific features. For example, some graphs grow over time according to a \emph{densification power law} which means that real graphs tend to sprout many more edges than nodes, and thus are densifying as they grow. Also the effective diameter of graphs tends to shrink or stabilize as the graph grows with time.

A related problem is \emph{data versioning} and its respective ability to query across multiple versions of data or to carry out general analysis.
%Considering graph
%data this problem is also common, for example  in the area of Linked
This problem has been investigated for instance within the domain of Linked
Open Data~\cite{DBLP:conf/semweb/Papakonstantinou16,DBLP:conf/esws/MeimarisP16,fernandez2015towards,fernandez2015bear}.

The respective data generator should hence be able to simulate a natural growth and/or changes in the structure of the graph with regards to the various features of distinct use cases. However,  even though the area of dynamic graphs is intensively studied, surprisingly there seem to exist only very few proposals of a generator for dealing with this area. In~\cite{GoerkeKlugeSchumm2012_1000029825} the authors focus on \emph{clustering dynamic graphs}, i.e. graphs where the clustering corresponds to the partition of nodes into natural groups based on the paradigm of intra-cluster density versus inter-cluster sparsity of edges. The generator generates a time series of random graphs $G_0, G_1, ..., G_n$, where $G_t$ emerges from $G_{t-1}$ via atomic updates, i.e., insertion or deletion of an edge or a vertex. The generator keeps track of a (dynamic) ground truth clustering. The probability of atomic events is chosen in a way that adheres to this clustering, without losing randomness.

@@

\subsection{Multi-model Data}
With the dawn of Big Data and especially its variety aspect, new types of
database management systems have emerged. One of the most interesting ones
are the so-called \emph{multi-model databases} that enable to store and
thus query across structurally different data. There exist various types of
multi-model systems combining  distinct subsets of Big Data structures including graph data.
For example, OrientDB\footnote{\url{http://orientdb.com/orientdb/}} which was
implemented on the basis of an object DBMS currently supports graph, document,
key/value, and object models. Such type of DBMSs also needs a specific
data generator that would enable
to test new features and analyze efficiency of operations. However, since the
multi-model systems are in the context of Big Data are rather new, there exist only
a few benchmarks targeting multi-model DBMSs (such as
Bigframe~\cite{journals/pvldb/KunjirKB14} or UniBench~\cite{conf/cidr/lu17})
with limited capabilities.

### 2018-09-24 comparison.tex
Adding TODOs.
Minor corrections
Sherif Changes

In order to provide the reader with a quick preview of the generators and thus to enable finding the target solutions easily, we start the survey with classification and comparative study of the existing tools. In general, there are various ways how to classify them. We first overview approaches used in related work and then we introduce our approach. As mentioned in the Introduction, since this survey is unique in terms of scope and new tools related to Big Data processing, our classification and comparative strategy differs too.

The graph data generators can be classified according to various criteria. For example, paper~\cite{DBLP:conf/sdm/ChakrabartiZF04} introduces two categories -- degree-based and procedural generators. Given a degree distribution (typically following a power law), \emph{degree-based generators} (e.g., Barabasi-Albert model~\cite{Barabasi99emergenceScaling}) try to find a graph that matches it, but without giving any insights about the graph or trying to match other criteria (like, e.g., small diameter, eigenvalues etc.). On the other hand, \emph{procedural generators} (e.g., R-MAT~\cite{DBLP:conf/sdm/ChakrabartiZF04}) try to find simple mechanisms to generate graphs that match a property of the real graphs and, typically, the power law degree distribution.

@@

@@

\begin{sidewaystable}
\scriptsize
\centering
\tbl{Key characteristics of the generators (part A)} {
\begin{tabular}{| c | p{2.2cm} | p{2cm} |  p{2.2cm} | l |  l | l | p{3cm} | p{1.4cm} | l | }
 \hline
           &   & \multicolumn{2}{c}{\textbf{Domain}}
               & \multicolumn{2}{|c|}{\textbf{Operations}}
               & \multicolumn{4}{c|}{\textbf{Configuration}}
               \\ \hline
           &  \textbf{Generator}
               & \textbf{Type}
               & \textbf{Target / sample}
               & \textbf{Read}
               & \textbf{Update}
               & \textbf{\rot{Properties}}
               & \textbf{Distributions}
			   & \textbf{Output}
               & \textbf{\rot{Distributed\ }}
               \\ \hline
\hline   % ====================== general graphs
\multirow{7}{*}{\rot{\textbf{General}}}
  & \cite{barabasi1999emergence} & fixed & & Y & N & Y &  & &  N  \\
\cline{2-10}
   & R-MAt & fixed & & Y & N & Y & power-law & &  N  \\
\cline{2-10}
  & GraphGen & & & & & & & &    \\
\cline{2-10}
  & Graph 500  & fixed & & Y & N & Y & power-law & &  Y  \\
\cline{2-10}
  & BTER & & & & & & & &    \\
\cline{2-10}
  & Darwini & & & & & & & &    \\
%  & Watts-Strogatz~\cite{watts1998collective} & fixed & & Y & N & Y &  & &  N  \\
%\cline{2-10}
\hline
\hline %  ======================  Semantic web
\multirow{20}{*}{\rot{\textbf{Semantic web}}}
 & LUBM & fixed & university  & fixed & N & Y & random (LCG) &  RDF & N   \\
\cline{2-10}
 & LBBM & extracted & Lehigh University BibTeX  & N & N & Y & Monte Carlo &  RDF & N   \\
\cline{2-10}
 & UOBM & fixed & university  & fixed & N & Y & random &  RDF & N   \\
\cline{2-10}
 & IIMB & fixed & movies  & N & N & Y & random &  RDF & N   \\
\cline{2-10}
 & BSBM & fixed & e-commerce  & fixed & N & Y & mostly normal &  RDF, relational & N   \\
\cline{2-10}
 & SP$^2$Bench & fixed & DBLP  & fixed & N & Y & based on DBLP  & RDF & N   \\
\cline{2-10}
 & \cite{Duan:2011:AOC:1989323.1989340} & extracted & -- & N & Y &N & -- &  RDF & N    \\
\cline{2-10}
 & DBPSB & extracted & DBpedia &  templates & N & Y & random &  RDF & N   \\
\cline{2-10}
 & LODIB & fixed & e-commerce &  N & N & Y & 44 types &  RDF & N   \\
\cline{2-10}
 & SIB & fixed & social network &  mixes & mix & Y & from real-world data &  RDF & N   \\
\cline{2-10}
 & Geographica & fixed & OpenStreetMap  & fixed + templates  & N & Y & -- &  RDF & N   \\
\cline{2-10}
 & WatDiv & schema-driven & user-defined  & templates & N & Y & uniform, normal, Zipfian &  RDF & N   \\
\cline{2-10}
 & RBench & extracted & DBLP, Yago  & templates & N & Y & from real-world data &  RDF & N  \\
\cline{2-10}
 & LDBC SPB & fixed & media  & mixes & N & Y & power law, skewed values, value correlation &  RDF & N  \\
\cline{2-10}
 & LinkGen & & & & & & & & \\
\hline
\end{tabular} }
\label{tab:comparisonCharacteristicsA}
\end{sidewaystable}

\begin{sidewaystable}
\scriptsize
\centering
\tbl{Key characteristics of the generators (part B)} {
\begin{tabular}{| c | p{2.2cm}| p{2cm} |  p{2.2cm} | l |  l | l | p{3cm} | p{1.4cm} | l | }
 \hline
           &   & \multicolumn{2}{c}{\textbf{Domain}}
               & \multicolumn{2}{|c|}{\textbf{Operations}}
               & \multicolumn{4}{c|}{\textbf{Configuration}}
               \\ \hline
           &  \textbf{Generator}
               & \textbf{Type}
               & \textbf{Target / sample}
               & \textbf{Read}
               & \textbf{Update}
               & \textbf{\rot{Properties}}
               & \textbf{Distributions}
			   & \textbf{Output}
               & \textbf{\rot{Distributed\ }}
               \\ \hline
\hline  % ======================  graph databases
\multirow{7}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & fixed  & social network  & fixed & Y & Y & power-law &  MAG &  Y  \\
\cline{2-10}
  & gMark & schema-driven (internal schema) &  user-defined  & generated &  N  & Y & uniform, normal, zipfian &  N-triples & N    \\
\cline{2-10}
  & graphGen & pattern-driven (Cypher schema) & user-defined  & -- & -- & Y & -- &  property graphs & N   \\
\hline
\hline %  ======================   social networks
\multirow{8}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & & & & & & & &    \\
\cline{2-10}
 & \cite{Yao2011} & & & & & & & &    \\
\cline{2-10}
 & LinkBench & & & & & & & &    \\
\cline{2-10}
 & S3G2 & Y & Y  & N &  & Y & Facebook-like  &  CSV, RDF & Y   \\
\cline{2-10}
 & \cite{Sukthankar-SocialInfo2014} & & & & & & & &    \\
\cline{2-10}
 & LDBC SNB  & Y & Y &   generated &  & Y & Facebook-like &  CSV, RDF & Y     \\
\cline{2-10}
  & \cite{Nettleton2016} & & & & & & & &    \\
%\cline{2-10}
% & SNG  & Y & Y &   N &  & Y & power-law & ? &  ?   \\
%\cline{2-10}
% & Forest Fire  & N & Y   & N &  & N & power-law &  ? & N   \\
%\cline{2-10}
% & Darwini  & N & N &   N &  & N & user-defined &  ? & Y   \\
%\cline{2-10}
% & BTER  & N & N &   N &  & N & user-defined &  ? & Y  \\
\hline
\hline   % ======================   graph analytics
\multirow{3}{*}{\rot{\textbf{Ana.}}}
  & HPC Scalable Graph Analysis & & & & & & & &    \\
\cline{2-10}
  & Graphalytics & & & & & & & &    \\
\hline
\hline   %  ======================  streaming
\multirow{2}{*}{\rot{\textbf{St.}}}
  & S2Gen & schema-driven & Social Network & Y & N & N & user-defined & RDF & N     \\
\cline{2-10}
  & RSPLab & schema-driven & Agnostic & Y & Y & N & user-defined & RDF & N     \\
%\cline{2-10}
%  & & & & & & & &  &   \\
\hline
\end{tabular} }
\label{tab:comparisonCharacteristicsB}
\end{sidewaystable}

@@

As we have mentioned, the basic classification we use in this paper is  based on a relatively vague division of the generators on the basis of the current application domains or research areas. In addition, some of the generators are either general or have features applicable in other domains. So the classes can overlap, as depicted in Table~\ref{tab:overlapping}.

\begin{table}[h]
\scriptsize
\centering
\tbl{Overlapping of classes of generators} {
\begin{tabular}{| c | l | l | l | l | l | l | l | }
 \hline
           &  \textbf{Generator}
               & \textbf{\rot{General}}
               & \textbf{\rot{Semantic Web}}
               & \textbf{\rot{Graph databases\ }}
               & \textbf{\rot{Social networks}}
               & \textbf{\rot{Analytics}}
               & \textbf{\rot{Steaming}}
               \\ \hline
\hline   % ====================== general graphs
\multirow{6}{*}{\rot{\textbf{General}}}
  & \cite{barabasi1999emergence} & x & & & & & \\
\cline{2-8}
   & R-MAt    & x & & & & & \\
\cline{2-8}
  & GraphGen  & x & & & & & \\
\cline{2-8}
  & Graph 500 & x & & & & & \\
\cline{2-8}
  & BTER      & x & & & & & \\
\cline{2-8}
  & Darwini   & x & & & & & \\
%  & Watts-Strogatz~\cite{watts1998collective}  & x & & & & & \\
%\cline{2-8}
\hline
\hline %  ======================  Semantic web
\multirow{15}{*}{\rot{\textbf{Semantic web}}}
 & LUBM  & & x & & & & \\
\cline{2-8}
 & LBBM  & & x & & & & \\
\cline{2-8}
 & UOBM  & & x & & & & \\
\cline{2-8}
 & IIMB & & x & & & & \\
\cline{2-8}
 & BSBM & & x & & & & \\
\cline{2-8}
 & SP$^2$Bench & & x & & & & \\
\cline{2-8}
 & \cite{Duan:2011:AOC:1989323.1989340} & & x & & & & \\
\cline{2-8}
 & DBPSB & & x & & & & \\
\cline{2-8}
 & LODIB & & x & & & & \\
\cline{2-8}
 & SIB & & x & & & & \\
\cline{2-8}
 & Geographica & & x & & & & \\
\cline{2-8}
 & WatDiv & & x & & & & \\
\cline{2-8}
 & RBench & & x & & & & \\
\cline{2-8}
 & LDBC SPB & & x & & & & \\
\cline{2-8}
 & LinkGen & & x & & & & \\
\hline
\hline  % ======================  graph databases
\multirow{3}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & & & x & & & \\
\cline{2-8}
  & gMark & & & x & & & \\
\cline{2-8}
  & graphGen & & & x & & & \\
\hline
\hline %  ======================   social networks
\multirow{7}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & & & & x & & \\
\cline{2-8}
 & \cite{Yao2011}  & & & & x & & \\
\cline{2-8}
 & LinkBench  & & & & x & & \\
\cline{2-8}
 & S3G2  & & & & x & & \\
\cline{2-8}
 & \cite{Sukthankar-SocialInfo2014}  & & & & x & & \\
\cline{2-8}
 & LDBC SNB   & & & & x & & \\
\cline{2-8}
  & \cite{Nettleton2016}  & & & & x & & \\
%\cline{2-8}
% & SNG  & & & & x & & \\
%\cline{2-8}
% & Forest Fire  & & & & x & & \\
%\cline{2-8}
% & Darwini & & & & x & & \\
%\cline{2-8}
% & BTER  & & & & x & & \\
\hline
\hline   % ======================   graph analytics
\multirow{2}{*}{\rot{\textbf{An.}}}
  & HPC Scalable Graph Analysis  & & & & & x & \\
\cline{2-8}
  & Graphalytics & & & & & x & \\
\hline
\hline   %  ======================  streaming
\multirow{2}{*}{\rot{\textbf{St.}}}
  & S2Gen & & & & & & x\\
\cline{2-8}
  & RSPLab & & & & & & x \\
%\cline{2-8}
%  & & & & & & \\
\hline
\end{tabular} }
\label{tab:overlapping}
\end{table}

>>>>

In order to provide the reader with a quick preview of the generators and thus to enable finding the target solutions easily, we start the survey with a  classification and comparative study of the existing tools. In general, there are various ways to classify them. We first provide an overview of the approaches used in related work after which we then introduce our approach. As mentioned in the Introduction, since this survey is unique in terms of scope and new tools related to Big Data processing, our classification and comparative strategy would differ as well.

The graph data generators can be classified according to various criteria. For example,~\cite{DBLP:conf/sdm/ChakrabartiZF04} introduces two categories -- degree-based and procedural generators. Given a degree distribution (typically following a power law), \emph{degree-based generators} (e.g., Barabasi-Albert model~\cite{Barabasi99emergenceScaling}) try to find a graph that matches it, but without giving any insights about the graph or trying to match other criteria (like, e.g., small diameter, eigenvalues etc.). On the other hand, \emph{procedural generators} (e.g., R-MAT~\cite{DBLP:conf/sdm/ChakrabartiZF04}) try to find simple mechanisms to generate graphs that match a property of the real graphs and, typically, the power law degree distribution.

@@

\TODO{Add info to all three tables.}

@@

\begin{sidewaystable}
\scriptsize
\centering
\tbl{Key characteristics of the generators (part A)} {
\begin{tabular}{| c | p{2.2cm} | p{2cm} |  p{2.2cm} | l |  l | l | p{3cm} | p{1.4cm} | l | }
 \hline
           &   & \multicolumn{2}{c}{\textbf{Domain}}
               & \multicolumn{2}{|c|}{\textbf{Operations}}
               & \multicolumn{4}{c|}{\textbf{Configuration}}
               \\ \hline
           &  \textbf{Generator}
               & \textbf{Type}
               & \textbf{Target / sample}
               & \textbf{Read}
               & \textbf{Update}
               & \textbf{\rot{Properties}}
               & \textbf{Distributions}
			   & \textbf{Output}
               & \textbf{\rot{Distributed\ }}
               \\ \hline
\hline   % ====================== general graphs
\multirow{7}{*}{\rot{\textbf{General}}}
  & \cite{barabasi1999emergence} & fixed & & Y & N & Y &  & &  N  \\
\cline{2-10}
   & R-MAt & fixed & & Y & N & Y & power-law & &  N  \\
\cline{2-10}
  & GraphGen & & & & & & & &    \\
\cline{2-10}
  & Graph 500  & fixed & & Y & N & Y & power-law & &  Y  \\
\cline{2-10}
  & BTER  & N & N &   N &  & N & user-defined &  ? & Y  \\
\cline{2-10}
  & Darwini  & N & N &   N &  & N & user-defined &  ? & Y   \\
%  & Watts-Strogatz~\cite{watts1998collective} & fixed & & Y & N & Y &  & &  N  \\
%\cline{2-10}
\hline
\hline %  ======================  Semantic web
\multirow{20}{*}{\rot{\textbf{Semantic web}}}
 & LUBM & fixed & university  & fixed & N & Y & random (LCG) &  RDF & N   \\
\cline{2-10}
 & LBBM & extracted & Lehigh University BibTeX  & N & N & Y & Monte Carlo &  RDF & N   \\
\cline{2-10}
 & UOBM & fixed & university  & fixed & N & Y & random &  RDF & N   \\
\cline{2-10}
 & IIMB & fixed & movies  & N & N & Y & random &  RDF & N   \\
\cline{2-10}
 & BSBM & fixed & e-commerce  & fixed & N & Y & mostly normal &  RDF, relational & N   \\
\cline{2-10}
 & SP$^2$Bench & fixed & DBLP  & fixed & N & Y & based on DBLP  & RDF & N   \\
\cline{2-10}
 & \cite{Duan:2011:AOC:1989323.1989340} & extracted & -- & N & Y &N & -- &  RDF & N    \\
\cline{2-10}
 & DBPSB & extracted & DBpedia &  templates & N & Y & random &  RDF & N   \\
\cline{2-10}
 & LODIB & fixed & e-commerce &  N & N & Y & 44 types &  RDF & N   \\
\cline{2-10}
 & SIB & fixed & social network &  mixes & mix & Y & from real-world data &  RDF & N   \\
\cline{2-10}
 & Geographica & fixed & OpenStreetMap  & fixed + templates  & N & Y & -- &  RDF & N   \\
\cline{2-10}
 & WatDiv & schema-driven & user-defined  & templates & N & Y & uniform, normal, Zipfian &  RDF & N   \\
\cline{2-10}
 & RBench & extracted & DBLP, Yago  & templates & N & Y & from real-world data &  RDF & N  \\
\cline{2-10}
 & LDBC SPB & fixed & media  & mixes & N & Y & power law, skewed values, value correlation &  RDF & N  \\
\cline{2-10}
 & LinkGen & & & & & & & & \\
\hline
\end{tabular} }
\label{tab:comparisonCharacteristicsA}
\end{sidewaystable}

\begin{sidewaystable}
\scriptsize
\centering
\tbl{Key characteristics of the generators (part B)} {
\begin{tabular}{| c | p{2.2cm}| p{2cm} |  p{2.2cm} | l |  l | l | p{3cm} | p{1.4cm} | l | }
 \hline
           &   & \multicolumn{2}{c}{\textbf{Domain}}
               & \multicolumn{2}{|c|}{\textbf{Operations}}
               & \multicolumn{4}{c|}{\textbf{Configuration}}
               \\ \hline
           &  \textbf{Generator}
               & \textbf{Type}
               & \textbf{Target / sample}
               & \textbf{Read}
               & \textbf{Update}
               & \textbf{\rot{Properties}}
               & \textbf{Distributions}
			   & \textbf{Output}
               & \textbf{\rot{Distributed\ }}
               \\ \hline
\hline  % ======================  graph databases
\multirow{7}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & fixed  & social network  & fixed & Y & Y & power-law &  MAG &  Y  \\
\cline{2-10}
  & gMark & schema-driven (internal schema) &  user-defined  & generated &  N  & Y & uniform, normal, zipfian &  N-triples & N    \\
\cline{2-10}
  & graphGen & pattern-driven (Cypher schema) & user-defined  & -- & -- & Y & -- &  property graphs & N   \\
\hline
\hline %  ======================   social networks
\multirow{8}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & & & & & & & &    \\
\cline{2-10}
 & \cite{Yao2011} & & & & & & & &    \\
\cline{2-10}
 & LinkBench & & & & & & & &    \\
\cline{2-10}
 & S3G2 & Y & Y  & N &  & Y & Facebook-like  &  CSV, RDF & Y   \\
\cline{2-10}
 & \cite{Sukthankar-SocialInfo2014} & & & & & & & &    \\
\cline{2-10}
 & LDBC SNB  & Y & Y &   generated &  & Y & Facebook-like &  CSV, RDF & Y     \\
\cline{2-10}
  & \cite{Nettleton2016} & & & & & & & &    \\
%\cline{2-10}
% & SNG  & Y & Y &   N &  & Y & power-law & ? &  ?   \\
%\cline{2-10}
% & Forest Fire  & N & Y   & N &  & N & power-law &  ? & N   \\
\hline
\hline   % ======================   graph analytics
\multirow{3}{*}{\rot{\textbf{Ana.}}}
  & HPC Scalable Graph Analysis & & & & & & & &    \\
\cline{2-10}
  & Graphalytics & & & & & & & &    \\
\hline
\hline   %  ======================  streaming
\multirow{2}{*}{\rot{\textbf{St.}}}
  & S2Gen & schema-driven & Social Network & Y & N & N & user-defined & RDF & N     \\
\cline{2-10}
  & RSPLab & schema-driven & Agnostic & Y & Y & N & user-defined & RDF & N     \\
%\cline{2-10}
%  & & & & & & & &  &   \\
\hline
\end{tabular} }
\label{tab:comparisonCharacteristicsB}
\end{sidewaystable}

As we can see ... \TODO{Add a summary.}

@@

As we have mentioned, the basic classification that we have used in this paper is  based on a relatively vague division of the generators on the basis of the current application domains or research areas. In addition, some of the generators are either general or have features applicable in other domains. So the classes can overlap, as depicted in Table~\ref{tab:overlapping}.

\begin{table}[h]
\scriptsize
\centering
\tbl{Overlapping of classes of generators} {
\begin{tabular}{| c | l | l | l | l | l | l | l | }
 \hline
           &  \textbf{Generator}
               & \textbf{\rot{General}}
               & \textbf{\rot{Semantic Web}}
               & \textbf{\rot{Graph databases\ }}
               & \textbf{\rot{Social networks}}
               & \textbf{\rot{Analytics}}
               & \textbf{\rot{Steaming}}
               \\ \hline
\hline   % ====================== general graphs
\multirow{6}{*}{\rot{\textbf{General}}}
  & \cite{barabasi1999emergence} & x & & & & & \\
\cline{2-8}
   & R-MAt    & x & & & & & \\
\cline{2-8}
  & GraphGen  & x & & & & & \\
\cline{2-8}
  & Graph 500 & x & & & & & \\
\cline{2-8}
  & BTER      & x & & & & & \\
\cline{2-8}
  & Darwini   & x & & & & & \\
\hline
\hline %  ======================  Semantic web
\multirow{15}{*}{\rot{\textbf{Semantic web}}}
 & LUBM  & & x & & & & \\
\cline{2-8}
 & LBBM  & & x & & & & \\
\cline{2-8}
 & UOBM  & & x & & & & \\
\cline{2-8}
 & IIMB & & x & & & & \\
\cline{2-8}
 & BSBM & & x & & & & \\
\cline{2-8}
 & SP$^2$Bench & & x & & & & \\
\cline{2-8}
 & \cite{Duan:2011:AOC:1989323.1989340} & & x & & & & \\
\cline{2-8}
 & DBPSB & & x & & & & \\
\cline{2-8}
 & LODIB & & x & & & & \\
\cline{2-8}
 & SIB & & x & & & & \\
\cline{2-8}
 & Geographica & & x & & & & \\
\cline{2-8}
 & WatDiv & & x & & & & \\
\cline{2-8}
 & RBench & & x & & & & \\
\cline{2-8}
 & LDBC SPB & & x & & & & \\
\cline{2-8}
 & LinkGen & & x & & & & \\
\hline
\hline  % ======================  graph databases
\multirow{3}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & & & x & & & \\
\cline{2-8}
  & gMark & & & x & & & \\
\cline{2-8}
  & graphGen & & & x & & & \\
\hline
\hline %  ======================   social networks
\multirow{7}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & & & & x & & \\
\cline{2-8}
 & \cite{Yao2011}  & & & & x & & \\
\cline{2-8}
 & LinkBench  & & & & x & & \\
\cline{2-8}
 & S3G2  & & & & x & & \\
\cline{2-8}
 & \cite{Sukthankar-SocialInfo2014}  & & & & x & & \\
\cline{2-8}
 & LDBC SNB   & & & & x & & \\
\cline{2-8}
  & \cite{Nettleton2016}  & & & & x & & \\
\hline
\hline   % ======================   graph analytics
\multirow{2}{*}{\rot{\textbf{An.}}}
  & HPC Scalable Graph Analysis  & & & & & x & \\
\cline{2-8}
  & Graphalytics & & & & & x & \\
\hline
\hline   %  ======================  streaming
\multirow{2}{*}{\rot{\textbf{St.}}}
  & S2Gen & & & & & & x\\
\cline{2-8}
  & RSPLab & & & & & & x \\
%\cline{2-8}
%  & & & & & & \\
\hline
\end{tabular} }
\label{tab:overlapping}
\end{table}

As we can see ... \TODO{Add a summary.}

### 2018-09-24 conclusion.tex
Sherif Changes

Graph data occur in a vast amount of distinct applications, such as biology, chemistry, physics, computer science, or social sciences, to name just a few. Graphs form one of the most complex data structures requiring specific and usually sophisticated approaches for processing and analysis. The history of graph theory, studying these structures and respective algorithms, can be traced back to the 18th century.

With the recent dawn of Big Data there have occurred also large scale graphs where the efficiency   of  processing methods is critical.  Approaches that work for smaller scale graphs often cannot be used, the data need to be processed in a distributed way and hence the efficiency is influenced by other aspects, such as limits of data transport. In addition, distribution of graphs, especially for highly connected cases, is a difficult task. Thus extensive testing of these methods for graphs of various sizes and structural complexity is extremely important.

>>>>

Graph data occur in a vast amount of distinct applications, such as biology, chemistry, physics, computer science, or social sciences, to name just a few. Graphs form one of the most complex data structures requiring specific and usually sophisticated approaches for processing and analysis. The history of graph theory,  that started from when these structures and their respective algorithms were studied can be traced back to the 18th century.

With the recent dawn of Big Data there have  been more occurrences of large scale graphs where the efficiency   of  processing methods is critical.  Approaches that work for smaller scale graphs often cannot be used, the data need to be processed in a distributed way and hence the efficiency is influenced by other aspects, such as limits of data transport. In addition, distribution of graphs, especially for highly connected cases, is a difficult task. Thus extensive testing of these methods for graphs of various sizes and structural complexity is extremely important.
